---
title: "ribor - R interface for .ribo files"
author: "Michael Geng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ribor - R interface for .ribo files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r global_options, include=FALSE}
library(knitr)
library(ribor)
```

# Introduction
This package is an R interface for .ribo files generated by the riboflow
pipeline. These .ribo files are HDF5 files, and it takes advantage of the
portability and group hierarchy capabilities offered by HDF5. The package
offers a suite of reading functions for each dataset and its respective 
metadata. Additionally, there are a few plotting functions that make use of 
the retrieved data.

Many of the functions use the `r Biocpkg("rhdf5")` package which is an 
interface for HDF5 in the **R** environment. This package makes use of data 
slicing,the act of accessing only a small subset of data at a time without 
reading in the entire dataset.

# Installation of the HDF5 package 

# Getting Started 

## Generating a ribo object and getting information

To get started, we should generate a ribo object that not only provides 
information about the file but most importantly a handle to the .ribo file
of interest.

```{r createRibo}
library(ribor)
file.path <- system.file("extdata", "sample.ribo", package = "ribor")
#generates the 'ribo' class object 
sample <- ribo(file.path)
```

The returned object is a list that contains 12 items.

Once we have called ribo object, we can then inquire about the contents of the
.ribo file.

First, we may want to look at general file information about the file and each
of its experiments.
```{r get_experiments}
#retrieves the experiments
sample.experiments <- get_experiments(ribo.object = sample)
```

Then, we may want to look at a list of experiment names, and then maybe look at the metadata of an experiment.
```{r get_info}
#retrieves the experiments
get_info(ribo.object = sample)

#get_metadata(ribo.object = sample,
#             "Hela_1")
```